---
title: "jfars"
author: "Jay Brown"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{jfars}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, warning=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(maps))
source("../R/fars_functions.R")
suppressMessages(library(knitr))
opts_knit$set(root.dir = "data/")

```

The Fatality Analysis Reporting System (FARS) is a database of car accidents where at least one person has died within 30 days of an accident taking place on a public road in the US. The data is used extensively to study traffic fatalities since 1975. 

The FARS data sample provides details of accidents from 2013 to 2015. This vignette reviews the functions available in this package.

## Dependencies

The FARS package depends on several standard R packages you may already have installed.

* dplyr
* tidyr
* readr 
* maps 


## Using the FARS Package

Use the FARS data package to explore incidents of traffic fatalities. The package will extract data from the data source noted and provide summary numbers by month and year as well as a map of a selected state indicating accident locations.

The public functions provide access to the data.

* fars_summarize_years 
* fars_map_state

### fars_summarize_years

The function fars_summarize_years reads available FARS data and will summarise it to provide a count of the number of fatal accidents per month in all states.

The function takes a vector of years and reads data from derived filenames then summarizes the number of observations for each year grouped by month. It transforms the data to wide format and returns this as a data frame.

#### Example

The example shows the number of accidents by month and by year.

```{r echo = TRUE}

fars_summarize_years(2013:2015)

```

### fars_map_state

The function fars_map_state takes the supplied state number and year and plots the locations of accidents on a map of the specified state. This is valuable to visualise the location of accidents within a particular state. The density of accidents could be used to plan highway and intersection improvements.

If an invalid state number is supplied it will error saying "invalid STATE number". State codes are standard codes as defined at [https://www2.census.gov/geo/docs/reference/state.txt]

If there are no accidents to plot, no map is plotted.

#### Example

The example shows the location of accidents in California in 2015.

```{r echo = TRUE}

fars_map_state(6, 2015)

```


## Private Functions

Private functions called from the above may be accessed using the convention jfars::fars_read. 

* fars_read Reads a specified file.
* make_filename Generates a filename by adding a supplied year to a filename.
* fars_read_years Reads multiple data files.
